---
title: "Homework Task 2 -- Basics in R"
date: 'Due: Thursday, May 2 by 08:59 CET'
author: "Moritz Mergheim, Hannah Feldmann"
output:
  html_document
---

# Instructions
* Work together within your group.
* Make sure you have R and RStudio installed. If you are an advanced user and aren't using RStudio, make sure you have `rmarkdown` in order to 'knit' the HTML output.
* Download [this zip file](https://michael-franke.github.io/XPLab_2019/homework/XPLab_2019_HW2.zip). It contains an .Rmd file that you will edit and submit.
* Open the .Rmd file in RStudio.
* Fill in your names in the 'author' heading.
* Fill in the required code and answers.
* 'Knit' the document (`ctrl/cmd` + `shift` + `K` in RStudio) to produce a HTML file.
* Push both the files to your group's Git repository before the deadline.
* Using Github Pages, make a page that displays the knitted html file.

# Suggested readings
* [R4DS](http://r4ds.had.co.nz/) (R for Data Science).

# Required R packages
* `tidyverse` (or `ggplot2`, `dplyr`, `purrr`, `tibble`)

```{r}
set.seed(123) # this ensures that any psuedo-random computations are always the same
```

___

# 1. Installing and running R

a.
---

Your first task is simply to show that you have been able to install
and run R and R Markdown.  You don't have to change this code, just
uncomment it. Then the correct output will automatically appear when
you 'knit' the document.


```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

R.version

sessionInfo()
```


**Which version of R are you running? On which platform are you running it?**

*ANSWER:*

**Version:** R version 3.5.3 (2019-03-11)

**Platform:** x86_64-w64-mingw32


b.
---

Install the package `tidyverse`. Don't install it in the code
below. Instead, install it through the console. Then write code below
to load the package and show the `sessionInfo` again.


```{r tidy = FALSE, message = TRUE, warning = TRUE, error = FALSE}

library(tidyverse)

sessionInfo()

```


**Which version of `tidyverse` do you have installed?**

*ANSWER:*

**Version:** tidyverse 1.2.1


___

# 2. Tibbles (tidy tables)


The ToothGrowth data set comes with base R. You can read about this data set
by running `?ToothGrowth` in the console. It is a data frame. In this
course, we prefer to use tibbles (tidy tables) instead of data frames.

a.
---

Convert the `ToothGrowth` data frame into a tibble using `as_tibble()`. Put
this in a new variable called `iris_tibble`. Then print the tibble
using the `print()` function.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

iris_tibble = as_tibble(ToothGrowth)
print(iris_tibble)


```



**Which data type is the variable "supp"?**

*ANSWER:*

**Data type:** "fct" hence a factor


b.
---

Starting from the complete `ToothGrowth` data set, filter only the guinea pigs
that were give orange juice.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

filter(ToothGrowth, supp == 'OJ')

```

**How many datapoints (i.e. guinea pigs) are left?**

*ANSWER:*

60-30=30



c.
---

Starting from the complete `ToothGrowth` data set, filter only the guinea pigs
that received ascorbic acid and had a dosage of 1.0mg or greater.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

filter(ToothGrowth, supp == 'VC', dose >= 1.0)



```



d.
---

Find out the mean length for the two supplement types. Do this with by
piping `tooth_growth_tibble` to `group_by()` and then to `summarise()`.  For
instructions read the help page for `summarise()`.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

ToothGrowth %>% group_by(supp) %>% summarise(mean_len = mean(len))


```


**What is the mean tooth length for guinea pigs that received Orange Juice?**

*ANSWER:*

**Mean length for OJ:** 20.7


___

# 3. Plotting data


a.
---

Using the `ToothGrowth` data set, create a plot of dose (x
axis) against length (y axis) using `ggplot()`. Show the two
supplement types in different colours.

```{r tidy = FALSE, message = FALSE, warning = FALSE, error = FALSE}

ToothGrowth %>% ggplot(aes(x = dose, y = len)) +
  geom_point() + 
  geom_point(data = filter(ToothGrowth, supp == 'OJ'), color = "orange", shape = "circle", size = 1) +
  geom_point(data = filter(ToothGrowth, supp == 'VC'), color = "yellow", shape = "circle", size = 1)


```


**After visually inspecting the graph, does there seem to be any difference between the types of supplement?**

*ANSWER:*

It seems like that for the lower doses the tooth length is higher for Guinea Pigs that received orange juice. For 2.0 mg the tooth length is mixed up.


___

End of homework sheet
